#list of tasks will do with an exisiting data base and R 

1. create logical and physical database design 
  entity-relationship diagrams are used to visualize data structures 


2. Clean and split the data into load files (usually used with programming languages - R or Python)



3. Create the physical database objects (can be done in or outside of R)
    use the logical databse to create the physical databse objects, such as tables 
```{r}
#creating the schools table using SQL in R 
library(RODBC)
df2 <- sqlQuery(con1, "Create Table
                DB2User1.school (
                B_ID CHAR(6) NOT NULL,
                S_ID CHAR(6) NOT NULL
                S_NAME VARCHAR(50),
                LEVEL VARCHAR(70),
                ENROLLMENT INTEGER)")

#dropping database tables
sqlDrop(con1, "DB2user1.school")
#clearing the databse table, keeps the structure in tact 
sqlClear(con1, "DB2user1.school")
```

4. Get the data into the databse (R)
```{r}
#The import command performs SQL insert statements for groups of rows and then commits the transaction periodically, not good for large datasets
import from board.csv
  of DEL
  message msg_board.txt
  replace into board 

#The load command, better option for large datasets, by passes logging mechanism
load from board.csv
  of del 
  replace into board
```

5. Analyze the data (R and SQL)




