#Wy use R with a relational database?
    - Because the data you want to work with resides in a relational database and R is a great data manipulation and visualization tool 
    - Because R has difficulty with manipulating large datasets, providing concurrent access to data, and (persistence)? 
          With R, data must reside in memory therefore SQL can be more efficiently to get it from database 
          SQL can be used to more efficiently extract subsets of a database within R
          Difficult to change observations within dataframes 
          
Relational Databases (RDBMSs) are optimized for:
    - Manipulating large structured databases 
    - providing concurrent data access to multiple users 
    - maintaining data consistency (data types/schemes) for all who want to pull data from it. Plus if you want to change an aspect about the database, the changes must adhere to specific parameters 
    - Persistence (granular, secure, consistent)
        R persistence options 
    
#Consider using character string for time or date representations, as the info is exchanged between R and the database (otherwise it'll guess what the date format should be)

```{r}
#Storing and retrieving R data structures 

saveRDS(df, file = "data.Rda")
df1 <- readRDS("data.Rda")
```
  
Database design considerations:
  - Relational databases minimized data redundancy (apply database normalization techniques, proper use of foreign keys)
      IN R, data redundancy must be handled with functions like unique() and duplicated() 
  - Model attributes and relationship accurately, enable data consistency (accurately model relationships between tables for data consistency )
      Declarative referential integrity: no equivalent options within dataframes in R 
  - Databse servers are efficient at merging data from different tables into a single structure 
      Instead of using the R merge() function, it's better to use SQL joins with virtual tables or views 
  - Databases support transactions and security (unit of work: antomicity, consistency, isolation, and durability - ACID properties)
      databse security protects database content, owners/users and prevents unauthorized access 


#Common database connectivity options for R: 
  -ODBC (open database connectivity): RODBC - open database connectivity using R, one of the most common R packages  
      Originally developed by Microsoft, can be used with many different database servers
          ODBC driver manager is platform specific (windows,linux)
          ODBC Drivers are database server specific (DB2, MySQL, Oracle)
          RODBC provides R functions to interact with database servers 
      
  -DBI-Based (Databse Interface-Based): 
      Database specific: RMySQL, RpostgreSQL, ROracle, RSQLite
      JDBC (Java database connectivity): RJDBC

```{r}
#Using RODBC

#Load the RODBC libary
library(RODBC)

#Connect to a database server 
con1 <- odbcConnect(dsn, userid, password)

#set a logical varaible to indicate if a connection was successful 
ds.connect <- (!is.null(con1) && con1 >= 0)
```